---
title: "Craft Beer Analysis"
author: "Chaoshun Hu, Mahesh Kuklani, Rene Pineda"
date: "June 25, 2018"
output:
  html_document:
    keep_md: true
---

# Craft Beer Market Analysis
###*Presented to the Brewers Association in Boulder, CO*

## **Section 1** Introduction

### The beer market in the U.S.

According to the Brewers Association, the overall beer market size in the U.S. was USD 111.4 Billion in 2017, of which USD 26 Billion belong to "craft beers". Retail dollar sales of craft increased 8%, a much larger growth than the rest of the industry. 

In volume terms, Overall U.S. beer volume sales were down 1% in 2017, whereas craft brewer sales continued to grow at a rate of 5% by volume, reaching 12.7% of the U.S. beer market by volume. Craft production grew the most for microbreweries.

We can therefore conclude that there is a surge in popularity of craft beers in the United States, produced by smaller companies. In contrast with the older and larger breweries that mostly produce lager beer, craft beer producers offer a wide variety of styles. Their manufacturing practices and quality of their products make these breweries more attractive to current consumers.

### Craft Breweries

The craft brewer definition is: An American craft brewer is small, independent, and traditional.

* Small: Annual production of 6 million barrels of beer or less (approximately 3 percent of U.S. annual sales). Beer production is attributed to a brewer according to the rules of alternating proprietorships.

* Independent: Less than 25 percent of the craft brewery is owned or controlled (or equivalent economic interest) by a beverage alcohol industry member which is not itself a craft brewer.

* Traditional: A brewer which has a majority of its total beverage alcohol volume in beers whose flavors derive from traditional or innovative brewing ingredients and their fermentation. Flavored Malt Beverages (FMBs) are not considered beers.

Due to the small size and explosive growth of craft breweries, the total number of breweries rose from 42 in 1978 to over 2,750 in 2012. This also produces a large amount of data, suitable to be analyzed with Data Science tools and techniques.

## **Section 2** Preparatory Steps

```{r preparatory steps, echo=TRUE, warning=FALSE, message=FALSE}
# Check the working directory
getwd()

#Rene's working directory
#setwd("E:/Bibliotecas/Documents/Data Science/SMU/MSDS 6306 Doing Data Science/Craft Beer Analysis/Rene Pineda Analysis")

#Mahesh's working directory
#setwd("E:/Mahesh/SMU/MSDS6306 Doing Data Science/Homework/Craft-Beer-Analysis")

#Chaoshun's working directory
#setwd("C:/Users/chux/Downloads/study1/study1")
setwd("/Users/chux/Desktop/datascience/Homework/Doing_Data_Science/study1")
#Set up your working directory here

#Load additional packages
library(dplyr)
library(ggplot2)
      
# Load the datasets
Beers <- read.csv("Beers.csv", header = TRUE)
Breweries <- read.csv("Breweries.csv", header = TRUE)

# Inspect and understand the structure of the datasets
str(Beers)
summary(Beers[,c(3,4,7)])
str(Breweries)
```
## **Section 3** Questions

### Question 1. How many breweries are present in each State?
```{r Question 1, echo = TRUE,warning=FALSE, message=FALSE}
table(Breweries$State, useNA = "no")
```
The distribution of breweries in each State can be seen in the following table:

State|Number of Breweries
-----|-------------------
AK   |7 
AL   |3
AR   |2
AZ   |11
CA   |39
CO   |47
CT   |8
DC   |1
DE   |2
FL   |15
GA   |7
HI   |4
IA   |5
ID   |5
IL   |18
IN   |22
KS   |3
KY   |4
LA   |5
MA   |23
MD   |7
ME   |9
MI   |32
MN   |12
MO   |9
MS   |2
MT   |9
NC   |19
ND   |1
NE   |5
NH   |3
NJ   |3
NM   |4
NV   |2
NY   |16
OH   |15
OK   |6
OR   |29
PA   |25
RI   |5
SC   |4
SD   |1
TN   |3
TX   |28
UT   |4
VA   |16
VT   |10
WA   |23
WI   |20
WV   |1
WY   |4

### Question 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r Question 2, echo=TRUE,warning=FALSE, message=FALSE}
# There is one "Name" variable in each dataset. Change Variable names to distinguish between Beer and Brewery names
# Each dataset have different Brewery ID variable name. Change to Brewery_ID to match both datasets:
Breweries <- rename(Breweries, Brewery_ID = Brew_ID, Brewery_Name = Name )
Beers <- rename(Beers, Brewery_ID = Brewery_id, Beer_Name = Name)

#Order the data by Brewery and Beer_ID (not necessary because we are using the merge command, but useful anyway)
Beers <- Beers[order(Beers$Brewery_ID,Beers$Beer_ID),]

#Check if the Brewery_ID values are the same in order to make sure they can be easily merged
unique(Beers$Brewery_ID) == unique(Breweries$Brewery_ID)

#Merge the datasets by Brewery_ID
MergedBeers <- merge(x = Beers, y = Breweries, by = "Brewery_ID", all = TRUE)
str(MergedBeers)

### Print the first 6 observations and the last six observations to check the merged file.
head(MergedBeers, 6)
tail(MergedBeers, 6)
```

### Question 3. Report the number of NA's in each column.
```{r Question 3, echo=TRUE,warning=FALSE, message=FALSE}
sapply(MergedBeers, function(y) sum(is.na(y)))
```
Two columns have NA's: ABV has 62 and IBU has 1005, out of 2410 total rows

### 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r Question 4, echo = TRUE,warning=FALSE, message=FALSE}
## First do not select data with NA as their mean would be NA then so use split MergedBeers on state and
## get the mean of the split values with na.rm = TRUE
ABV.median <- sapply(split(MergedBeers, MergedBeers$State), function(y) median(y$ABV, na.rm = TRUE))

IBU.median <- sapply(split(MergedBeers, MergedBeers$State), function(y) median(y$IBU, na.rm = TRUE))

## Sort it in descending order
ABV.median.sort <- ABV.median[order(-ABV.median)]

IBU.median.sort <- IBU.median[order(-IBU.median)]

##plot a bar chart to compare
##Barplot of ABV value(s) vs State
barplot(ABV.median.sort, main="ABV vs State", xlab="State", ylab="ABV Value(s)", las=2)
##Barplot of IBU value(s) vs State
barplot(IBU.median.sort, main="IBU vs State", xlab="State", ylab="IBU Value(s)", las=2)
```

### 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
#### State that has maximum alcoholic (ABV) beer is CO.  
#### State that has most bitter (IBU) beer is OR.  
```{r Question 5, echo = TRUE,warning=FALSE, message=FALSE}
##State with maximum alcoholic (ABV) beer
which(MergedBeers$ABV == max(MergedBeers$ABV, na.rm=TRUE))

MergedBeers$State[which(MergedBeers$ABV == max(MergedBeers$ABV, na.rm=T))]

## State with maximum alcoholic (IBU) beer
MergedBeers$State[which(MergedBeers$IBU == max(MergedBeers$IBU, na.rm=T))]
```

### 6. Summary statistics for the ABV variable.

```{r Question 6, echo = TRUE,warning=FALSE, message=FALSE}
##Sumary statistics for the ABV variable from the Beers dataset
summary(Beers$ABV)

## Summary statistics for the ABV variable of the merged dataset
summary(MergedBeers$ABV)
```

### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
#### There seems to be an apparent relationship between the bitterness of the beer and its alcoholic content  
```{r Question 7, echo = TRUE,warning=FALSE, message=FALSE}
##Scatter plot between bitterness of beer and its alcoholic content
plot(IBU~ABV, data=MergedBeers)
abline(lm(IBU~ABV, data=MergedBeers), col="red")

```


## **Section 4** Conclusions
In summary, this primary objective of this work is to take two different data files which are beer and breweries data, read the data from the a csv file into a data frame, inspect and understand the structure of the data, merge the data frames, and perform some analysis on the final data set.

As Data Scientist, it is very rare to work only on a single perfect data and thus a large percentage of work will be accept different datasets, merge different available data sets before processing it as illustrated in this work. After preparing the data, statistical inference can then be made to the data. 

Based on the analysis, California and Colorado are top two which have most breweries. If order the all the data by Brewery and Beer IDs. The first six observations include beer names: Parapet ESB, Stronghold, Pumpion, Wall's End, Maggie's Leap and Get Together. The last 6 obervations include beer names: Pilsner Ukiah, Porkslap Pale Ale, Moo Thunder Stout, Snapperhead IPA, Heinnieweisse Weissebier and Urban Wilderness Pale Ale; Out of 2410 total rows of the data, there are two columns having NA's: ABV has 62 and IBU has 1005. 

After computing the median alcohol content and international bitterness unit for each state. DC has the highest median ABV, MS has the highest median IBV; UT has the lowest median ABV, SD has the lowest median IBV. But Co has the mximum alcoholic (ABV) beer and OR has the most bitter(IBU) beer. Afer running the summary statistics to ABV variable, the min ABV is 0.001 and the max ABV is 0.128. The Mean is 0.05977 and the median is 0.056. It has a littble bit skewness here. Based on the scatter plot, there is no linear relation between the bitterness of the beer and the alcoholic cotent.

