---
title: "Craft Beer Analysis codeBook"
author: "Rene Pineda, Mahesh Kuklani, Chaoshun Hu"
date: "June 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Description

The Beers dataset contains a list of 2410 US craft beers and Breweries dataset contains 558 US breweries. The datasets descriptions are as follows.

## Beer data description

###report overview
The beer dataset examined has the following dimensions:


---------------------------------
Feature                    Result
------------------------ --------
Number of observations       2410

Number of variables             7
---------------------------------


### Beer data summary table

------------------------------------------------------------------------
Label   Variable            Class       # unique  Missing  Description  
                                          values                        
------- ------------------- --------- ---------- --------- -------------
        **[Name]**          factor          2305  0.00 %                

        **[Beer\_ID]**      integer         2410  0.00 %                

        **[ABV]**           numeric           75  2.57 %                

        **[IBU]**           integer          108  41.70 %               

        **[Brewery\_id]**   integer          558  0.00 %                

        **[Style]**         factor           100  0.00 %                

        **[Ounces]**        numeric            7  0.00 %                
------------------------------------------------------------------------




### Beer data variable list
#### Name


---------------------------------------------
Feature                                Result
------------------------- -------------------
Variable type                          factor

Number of missing obs.                0 (0 %)

Number of unique values                  2305

Mode                        "Nonstop Hef Hop"
---------------------------------------------

---

#### Beer\_ID


-------------------------------------------
Feature                              Result
------------------------- -----------------
Variable type                       integer

Number of missing obs.              0 (0 %)

Number of unique values                2410

Median                               1453.5

1st and 3rd quartiles       808.25; 2075.75

Min. and max.                       1; 2692
-------------------------------------------




---

#### ABV


---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                   numeric

Number of missing obs.      62 (2.57 %)

Number of unique values              74

Median                             0.06

1st and 3rd quartiles        0.05; 0.07

Min. and max.                   0; 0.13
---------------------------------------




---

#### IBU


-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                     integer

Number of missing obs.      1005 (41.7 %)

Number of unique values               107

Median                                 35

1st and 3rd quartiles              21; 64

Min. and max.                      4; 138
-----------------------------------------

---

#### Brewery\_id


-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               integer

Number of missing obs.      0 (0 %)

Number of unique values         558

Median                          206

1st and 3rd quartiles       94; 367

Min. and max.                1; 558
-----------------------------------




---

#### Style


------------------------------------------
Feature                             Result
------------------------- ----------------
Variable type                       factor

Number of missing obs.             0 (0 %)

Number of unique values                100

Mode                        "American IPA"
------------------------------------------




---

### Ounces


-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           7

Median                           12

1st and 3rd quartiles        12; 16

Min. and max.               8.4; 32
-----------------------------------

---

## Breweries data description

### report overview
The breweries dataset examined has the following dimensions:


---------------------------------
Feature                    Result
------------------------ --------
Number of observations        558

Number of variables             4
---------------------------------


### Breweries data summary table

---------------------------------------------------------------------
Label   Variable         Class       # unique  Missing  Description  
                                       values                        
------- ---------------- --------- ---------- --------- -------------
        **[Brew\_ID]**   integer          558  0.00 %                

        **[Name]**       factor           551  0.00 %                

        **[City]**       factor           384  0.00 %                

        **[State]**      factor            51  0.00 %                
---------------------------------------------------------------------




### Variable list

#### Brew ID


------------------------------------------
Feature                             Result
------------------------- ----------------
Variable type                      integer

Number of missing obs.             0 (0 %)

Number of unique values                558

Median                               279.5

1st and 3rd quartiles       140.25; 418.75

Min. and max.                       1; 558
------------------------------------------




---

#### Name


------------------------------------------------
Feature                                   Result
------------------------- ----------------------
Variable type                             factor

Number of missing obs.                   0 (0 %)

Number of unique values                      551

Mode                        "Blackrocks Brewery"
------------------------------------------------




---

#### City


--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                   factor

Number of missing obs.         0 (0 %)

Number of unique values            384

Mode                        "Portland"
--------------------------------------




---

#### State


-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values          51

Mode                          " CO"
-----------------------------------

---

## **Section 4** Preparatory Steps


```{r, echo=FALSE}
# Check the working directory
getwd()
```

```

```

```{r, echo=FALSE}
#Rene's working directory
setwd("E:/Bibliotecas/Documents/Data Science/SMU/MSDS 6306 Doing Data Science/Craft Beer Analysis")

#Mahesh's working directory
#setwd("E:/Mahesh/SMU/MSDS6306 Doing Data Science/Homework/Craft-Beer-Analysis")

#Chaoshun's working directory
#setwd("C:/Users/chux/Downloads/study1/study1")
#setwd("/Users/chux/Desktop/datascience/Homework/Doing_Data_Science/study1")
#Set up your working directory here

#Load additional packages
library(dplyr)
library(ggplot2)
      
# Load the datasets
Beers <- read.csv("Beers.csv", header = TRUE)
Breweries <- read.csv("Breweries.csv", header = TRUE)

Breweries <- rename(Breweries, Brewery_ID = Brew_ID, Brewery_Name = Name )
Beers <- rename(Beers, Brewery_ID = Brewery_id, Beer_Name = Name)

#Order the data by Brewery and Beer_ID (not necessary because we are using the merge command, but useful anyway)
Beers <- Beers[order(Beers$Brewery_ID,Beers$Beer_ID),]

#Merge the datasets by Brewery_ID
MergedBeers <- merge(x = Beers, y = Breweries, by = "Brewery_ID", all = TRUE)
```

## **Section 5** Questions

### Question 1. How many breweries are present in each State?
```{r Question 1, echo = TRUE,warning=FALSE, message=FALSE}
table(Breweries$State, useNA = "no")
```

### Question 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r, echo=FALSE}
head(MergedBeers, 6)
tail(MergedBeers, 6)
```

### Question 3. Report the number of NA's in each column.
```{r Question 3, echo=TRUE,warning=FALSE, message=FALSE}
sapply(MergedBeers, function(y) sum(is.na(y)))
```
Two columns have NA's: ABV has 62 and IBU has 1005, out of 2410 total rows

```

### 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r Question 4, echo = TRUE,warning=FALSE, message=FALSE}
## First do not select data with NA as their mean would be NA then so use split MergedBeers on state and
## get the mean of the split values with na.rm = TRUE
ABV.median <- sapply(split(MergedBeers, MergedBeers$State), function(y) median(y$ABV, na.rm = TRUE))

IBU.median <- sapply(split(MergedBeers, MergedBeers$State), function(y) median(y$IBU, na.rm = TRUE))

## Sort it in descending order
ABV.median.sort <- ABV.median[order(-ABV.median)]

IBU.median.sort <- IBU.median[order(-IBU.median)]

##plot a bar chart to compare
##Barplot of ABV value(s) vs State
barplot(ABV.median.sort, main="ABV vs State", xlab="State", ylab="ABV Value(s)", las=2)
##Barplot of IBU value(s) vs State
barplot(IBU.median.sort, main="IBU vs State", xlab="State", ylab="IBU Value(s)", las=2)
```

### 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
#### State that has maximum alcoholic (ABV) beer is CO.  
#### State that has most bitter (IBU) beer is OR.  
```{r Question 5, echo = TRUE,warning=FALSE, message=FALSE}
##State with maximum alcoholic (ABV) beer
which(MergedBeers$ABV == max(MergedBeers$ABV, na.rm=TRUE))

MergedBeers$State[which(MergedBeers$ABV == max(MergedBeers$ABV, na.rm=T))]

## State with maximum alcoholic (IBU) beer
MergedBeers$State[which(MergedBeers$IBU == max(MergedBeers$IBU, na.rm=T))]
```

### 6. Summary statistics for the ABV variable.

```{r Question 6, echo = TRUE,warning=FALSE, message=FALSE}
##Sumary statistics for the ABV variable from the Beers dataset
summary(Beers$ABV)

## Summary statistics for the ABV variable of the merged dataset
summary(MergedBeers$ABV)
```

### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
#### There seems to be an apparent relationship between the bitterness of the beer and its alcoholic content  
```{r Question 7, echo = TRUE,warning=FALSE, message=FALSE}
##Scatter plot between bitterness of beer and its alcoholic content
plot(IBU~ABV, data=MergedBeers)
abline(lm(IBU~ABV, data=MergedBeers), col="red")

```

